Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-04-05 Manuscript/manuscript.Rmd
editted manuscript

By isolating bins that overlapped with flanked TAD boundaries we were able to assess the correlations of genomic annotations and determine which annotations were enriched together at boundaries. To  get a hollisitic view of how some annotations tended to occur together at boundaries, we computed the pairwise correlations of the averages between count overlaps, count percents, and $log_{2}$ distances of each genomic annotation with bins that overlapped with flanked TAD boundaries. Clustered correlograms at the 10 kb resolution, for both cell lines, showed that distinct groups of transcription factors were highly correlated including CTCF, ZNF143, RAD21, and SMC3ab (Figure 7). Likewise, combinations of particular histone modifications and chromatin states were also shown to exhibit similar profiles near TAD boundaries including H3K4me1, H3K4me2, and H3K4me3, aswell as strong and weak enhancers. Furthermore, particular chromatin segmentation states, inclduing heterochromatin, repressed chromatin, poised promotors, and repetitive copy number variations, were found to be highly correlated, yet inversely correlated with many transcription factor binding sites. Similar correlations were exhibited across the remainder of resolutions, for both cell lines (Supplementary Figure X). The high correlations among many of the genomic annotations gave precedent for the use of an elastic-net regularization algorithm in order to suitably reduce the feature space while retaining possibly correlated annotations that could be predictive of TAD boundary formation. The regularization was performed predictor-type and resolution specific data after implementing each of the resampling techniques within each cell line.

@@

Lastly, we investigated which genomic annotations were most predictive of TAD boundary formation. Looking at heatmaps of variable importances provided by the random forest classifiers implemented on SMOTE re-sampled data, with distance-type predictors, we found that known transcription factors associated with the hypothesized loop extrusion model [??? include refs] were give high predictive importance in the model (Figure 10). These tended to include the SMC3ab, ZNF143, CTCF, and RAD21 transcription factor binding sites and were found to be ranked consistly across resolution. Other annotations known to be associated with gene regulation and affect structural properties of chromatin, such as the YY1 and MAZ transcription factors, as well DNAse hypersensitive sites, regions of heterochromatin, and the H3k27 trimethylation and H3K4me1 modifications were less consistently ranked across resolutions for either cell line.

@@

# LIMITATIONS

>>>>

By isolating bins that overlapped with flanked TAD boundaries we were able to assess the correlations of genomic annotations and determine which annotations were enriched together at boundaries. To  get a hollisitic view of how some annotations tended to occur together at boundaries, we computed the average pairwise correlations between count overlaps, count percents, and $log_{2}$ distances of each genomic annotation with bins that overlapped with flanked TAD boundaries. Clustered correlograms at the 10 kb resolution, for both cell lines, showed that distinct groups of transcription factors were highly correlated including CTCF, ZNF143, RAD21, and SMC3ab (Figure 7). Likewise, combinations of particular histone modifications and chromatin states were also shown to exhibit similar profiles near TAD boundaries including H3K4me1, H3K4me2, and H3K4me3, aswell as strong and weak enhancers. Furthermore, particular chromatin segmentation states, inclduing heterochromatin, repressed chromatin, poised promotors, and repetitive copy number variations, were found to be highly correlated, yet inversely correlated with many transcription factor binding sites. Similar correlations were exhibited across the remainder of resolutions, for both cell lines (Supplementary Figures 2-4). The high correlations among many of the genomic annotations gave precedent for the use of an elastic-net regularization algorithm in order to suitably reduce the feature space while retaining possibly correlated annotations that could be predictive of TAD boundary formation. The regularization was performed predictor-type and resolution specific data after implementing each of the resampling techniques within each cell line.

@@

Lastly, we investigated which genomic annotations were most predictive of TAD boundary formation. Looking at heatmaps of variable importances provided by the random forest classifiers implemented on SMOTE re-sampled data, with distance-type predictors, we found that known transcription factors associated with the hypothesized loop extrusion model [@fudenberg2016formation; @alipour2012self] were give high predictive importance in the model (Figure 10). These tended to include the SMC3ab, ZNF143, CTCF, and RAD21 transcription factor binding sites and were found to be ranked consistly across resolution. Other annotations known to be associated with gene regulation and affect structural properties of chromatin, such as the YY1 and MAZ transcription factors, as well DNAse hypersensitive sites, regions of heterochromatin, and the H3k27 trimethylation and H3K4me1 modifications were less consistently ranked across resolutions for either cell line.

@@

## LIMITATIONS
